---
title: "VNRP_data"
author: "Marc Peipoch"
date: "7/23/2021"
output: html_document
---

Will use old VNRP data from Warm Springs to Turah to help develop the core ideas of the paper regarding nutrient removal efficiency.
This is a simple calculation of the how Efficiency (calculated as in original Valett and Peipoch 2020 submission) relates to reach length in the UCFR. 

```{r setup, include=FALSE}
library(ggplot2); library(dplyr)
#download the data
urlfile<-'https://raw.githubusercontent.com/mpstroud/NPDefficiency/main/NPD_full_data.csv'
NPD_full_data<-read.csv(url(urlfile))

#general look at NO3eff versus reach length
ggplot(data= NPD_full_data, aes(L,NO3eff_total,color=season)) +
  geom_point(size=3) +
  theme_minimal() 
```

>We need to compare L-NO3eff incrementally, not simply by reach or the character of the reach dominates the relation.

```{r}
proba = NPD_full_data %>%
          filter(season == "3summer") %>%
          ggplot(., aes(L,NO3eff_total,color=season)) +
  geom_point(size=3) +
  theme_minimal() 

```
